Cultural Transmission and Technological Transitions during the Late Paleolithic in Korea

### 2023-02-08 analysis/paper/paper.qmd
editing correlation

One limitation of this aggregation of all stemmed points in our sample into each chronological phase is that the sample consists of a relatively large number of assemblages with only 1-2 stemmed points. These isolated finds are ambiguous with respect to a local tradition of artifact making, so to further investigate temporal change, we focus only the four assemblages that have more than five stemmed points (@fig-plot-cor-four-assemblage). We assume these four assemblages are more likely to represent a consistent, recurring way of making stemmed points than isolated finds, and thus more relevant for comparing modes of cultural transmission of artifact making. The four assemblages include three from the Suyanggae site (SYG) and one from the Yongsandong site (YS). There are more variation in correlation coefficient in Phase 2 while the median values for the two assemblages from Phase 3 are similar to each other and close to `r threshold_value_cor`. Even though SYG1_2 is from Phase 2 but it has the higher median than other assemblages.

@@

## Correlation between sites and raw materials

@fig-corr-raw-ass

```{r}
#| label: fig-corr-raw-ass
#| fig-cap: A. Correlation for each site that has more than five stemmed points. B. Correlation for raw materials that were used for making more than one stemmed poitns. The grey points represent correlation between two attributes.

@@

# correlation by site
df_sitename_corr_site <- 
  df_sitename %>%
  group_by(sitename) %>% 
  filter(n() > 5) %>% 
  dplyr::select(-sitename, 
                -SPstage.Stage) %>% 
  nest(-sitename) %>% 
  mutate(corrr = purrr::map(data, ~correlate(.x) %>%  
                                   shave() %>% 
                                  pivot_longer(-term))) %>% 
  unnest(corrr) %>% 
  filter()

correlations_by_site <- 
  ggplot(df_sitename_corr_site) +
  aes(sitename,
      value) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2,
                               alpha = 0.3) +
  theme_minimal() +
  labs(y = "Correlation coefficient",
       x = "Site") +
theme(axis.title.y = element_text(size = 4)) +
  # draw a line to show our threshold value
  geom_hline(yintercept = threshold_value_cor, colour = "red") +
  theme_bw()

# correlation by raw material
df_sitename_corr_raw_material <- 
  df_sitename %>%
  group_by(SPstage.Raw_material) %>% 
  filter(n() > 2) %>% 
  dplyr::select(-sitename, 
                -SPstage.Stage) %>% 
  nest(-SPstage.Raw_material) %>% 
  mutate(corrr = purrr::map(data, ~correlate(.x) %>%  
                                   shave() %>% 
                                  pivot_longer(-term))) %>% 
  unnest(corrr) %>% 
  filter()

correlations_by_raw_material <- 
  ggplot(df_sitename_corr_raw_material) +
  aes(SPstage.Raw_material,
      value) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2,
                               alpha = 0.3) +
  theme_minimal() +
  labs(y = "Correlation coefficient",
       x = "Raw material") +
theme(axis.title.y = element_text(size = 4)) +
  # draw a line to show our threshold value
  geom_hline(yintercept = threshold_value_cor, colour = "red") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  theme_bw()

library(cowplot)
plot_grid(correlations_by_site,
          correlations_by_raw_material,
                    ncol = 1,
                    labels = c("A", "B"))

@@

<!-- BM: overall I think the CV values are really not that interesting or informative. They are difficult to interpret because they all cluster around 15-30, which seems to me to be the midpoint between 'highly standardized' (3-5) and 'highly different' (40-50). So I think we will struggle to convince any readers that we have any strong signals of cultural transmission biases in the CV data. Probably we can satisfy most readers that guided variation is the main bias in these data. That's my perspective after reading this paper several times, and re-reading several related papers by Bettiner and Eerkens and Lipo, and others  -->

>>>>

```
## Correlation between assemblages

One limitation of this aggregation of all stemmed points in our sample into each chronological phase is that the sample consists of a relatively large number of assemblages with only 1-2 stemmed points. These isolated finds are ambiguous with respect to a local tradition of artifact making, so to further investigate temporal change, we focus only the four assemblages that have more than five stemmed points (@fig-plot-cor-four-assemblage). We assume these four assemblages are more likely to represent a consistent, recurring way of making stemmed points than isolated finds, and thus more relevant for comparing modes of cultural transmission of artifact making.The four assemblages include three from the Suyanggae site (SYG) and one from the Yongsandong site (YS). There are more variation in correlation coefficient in Phase 2 while the median values for the two assemblages from Phase 3 are similar to each other and close to `r threshold_value_cor`. There is no clear pattern between assemblages. SYG1_2 is from Phase 2 but it has the higher median than the assemblage from the same site in the later phase.

@@

## Correlation between raw materials

To examine the impact of raw materials on stemmed points, we computed the correlation for different raw materials. We excluded raw materials that were used for less than three points. @fig-corr-raw-ass shows that stemmed points made out of acidic volcanic rocks are highly correlated between most of the attributes but some of the attributes are negatively related. Correlation values for porphyry stemmed points have a narrow distribution. Rhyolite stemmed points have widely distributed correlations. Some of the attributes are negatively related. Like acidic volcanic rocks, the correlation of shale stemmed points is also skewed into two groups. Therefore, we assume that the correlation varies depending on raw materials.

```{r}
#| label: fig-corr-raw-ass
#| fig-cap: Correlation for raw materials that were used for making more than one stemmed poitns. The grey points represent correlation between two attributes.

@@

# correlation by site
# df_sitename_corr_site <- 
#   df_sitename %>%
#   group_by(sitename) %>% 
#   filter(n() > 5) %>% 
#   dplyr::select(-sitename, 
#                 -SPstage.Stage) %>% 
#   nest(-sitename) %>% 
#   mutate(corrr = purrr::map(data, ~correlate(.x) %>%  
#                                    shave() %>% 
#                                   pivot_longer(-term))) %>% 
#   unnest(corrr) %>% 
#   filter()
# 
# correlations_by_site <- 
#   ggplot(df_sitename_corr_site) +
#   aes(sitename,
#       value) +
#   geom_boxplot() +
#   ggbeeswarm::geom_quasirandom(size = 2,
#                                alpha = 0.3) +
#   theme_minimal() +
#   labs(y = "Correlation coefficient",
#        x = "Site") +
# theme(axis.title.y = element_text(size = 4)) +
#   # draw a line to show our threshold value
#   geom_hline(yintercept = threshold_value_cor, colour = "red") +
#   theme_bw()

# correlation by raw material
df_sitename_corr_raw_material <- 
  df_sitename %>%
  group_by(SPstage.Raw_material) %>% 
  filter(n() > 3) %>% 
  dplyr::select(-sitename, 
                -SPstage.Stage) %>% 
  nest(-SPstage.Raw_material) %>% 
  mutate(corrr = purrr::map(data, ~correlate(.x) %>%  
                                   shave() %>% 
                                  pivot_longer(-term))) %>% 
  unnest(corrr) %>% 
  filter()

# correlations_by_raw_material <- 
  ggplot(df_sitename_corr_raw_material) +
  aes(SPstage.Raw_material,
      value) +
  geom_boxplot() +
  ggbeeswarm::geom_quasirandom(size = 2,
                               alpha = 0.3) +
  theme_minimal() +
  labs(y = "Correlation coefficient",
       x = "Raw material") +
theme(axis.title.y = element_text(size = 4)) +
  # draw a line to show our threshold value
  geom_hline(yintercept = threshold_value_cor, colour = "red") +
  # scale_x_discrete(guide = guide_axis(n.dodge = 2))+
  theme_bw()

@@
